---
title: "Project 1"
author: "Yuhan Xiao"
output: html_document
date: "2023-09-16"
---

## Background

```{r message=FALSE, warning=FALSE}
library(here)
library(tidyverse)
library(dplyr)

# tests if a directory named "data" exists locally
if (!dir.exists(here("data"))) {
    dir.create(here("data"))
}

# saves data only once (not each time you knit a R Markdown)
if (!file.exists(here("data", "chocolate.RDS"))) {
    url_csv <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-18/chocolate.csv"
    chocolate <- readr::read_csv(url_csv)

    # save the file to RDS objects
    saveRDS(chocolate, file = here("data", "chocolate.RDS"))
}

# read the data locally
chocolate <- readRDS(here("data", "chocolate.RDS"))
as_tibble(chocolate)

# take a glimpse of the data
glimpse(chocolate)
```

## Part 1: Explore data

1. Make a histogram of the rating scores to visualize the overall distribution of scores. Change the number of bins from the default to 10, 15, 20, and 25. Pick on the one that you think looks the best. Explain what the difference is when you change the number of bins and explain why you picked the one you did.

As I increased the number of bins, some bins became empty due to lacking values within that range. I picked 10 bins because no bin is empty in this case.

```{r}
hist(chocolate$rating, 
     breaks=seq(1,4,3/10),
     xlab="Rating", ylab="Frequency", 
     main="Frequency of Chocolate Rating")
```

2. Consider the countries where the beans originated from. How many reviews come from each country of bean origin?
```{r}
chocolate %>%
  select(rating,country_of_bean_origin) %>%
  group_by(country_of_bean_origin) %>%
  summarise(
    number_of_reviews = n()
  )
```

3. What is average rating scores from reviews of chocolate bars that have Ecuador as country_of_bean_origin in this dataset? For this same set of reviews, also calculate (1) the total number of reviews and (2) the standard deviation of the rating scores. Your answer should be a new data frame with these three summary statistics in three columns. Label the name of these columns mean, sd, and total.

```{r}
summary_data <- chocolate %>%
  select(rating,country_of_bean_origin) %>%
  filter(country_of_bean_origin == "Ecuador") %>%
  summarise(
    mean = mean(rating),
    sd = sd(rating),
    total = n()
  )
summary_data
```

4. Which company (name) makes the best chocolate (or has the highest ratings on average) with beans from Ecuador?

```{r}
chocolate %>%
  select(rating,country_of_bean_origin,company_manufacturer) %>%
  filter(country_of_bean_origin == "Ecuador") %>%
  group_by(company_manufacturer) %>%
  summarise(
    mean = mean(rating),
  ) %>%
  arrange(desc(mean)) %>%
  head(5)
```

5. Calculate the average rating across all country of origins for beans. Which top 3 countries (for bean origin) have the highest ratings on average?

```{r}
chocolate %>%
  select(rating,country_of_bean_origin) %>%
  group_by(country_of_bean_origin) %>%
  summarise(
    mean = mean(rating),
  ) %>%
  arrange(desc(mean)) %>%
  head(3)
```

6. Following up on the previous problem, now remove any countries of bean origins that have less than 10 chocolate bar reviews. Now, which top 3 countries have the highest ratings on average?

```{r}
chocolate %>%
  select(rating,country_of_bean_origin) %>%
  group_by(country_of_bean_origin) %>%
  summarise(
    mean = mean(rating),
    total = n()
  ) %>%
  filter(total >= 10) %>%
  arrange(desc(mean)) %>%
  head(3)
```

7. On average, which category of chocolate percentage is most highly rated? Do these countries mostly agree or are there disagreements?

Chocolates with cocoa percentage between 60% to 70% have the highest ratings on average. 
  
```{r}
# first step: Identify the countries of bean origin with at least 50 reviews. Remove reviews from countries are not in this list.
countries_with_enough_reviews <- chocolate %>%
  select(country_of_bean_origin) %>%
  group_by(country_of_bean_origin) %>%
  summarise(
    total = n()
  ) %>%
  filter(total >= 50) %>%
  pull(country_of_bean_origin)

# second step: Using the variable describing the chocolate percentage for each review, create a new column that groups chocolate percentages into one of four groups: (i) <60%, (ii) >=60 to <70%, (iii) >=70 to <90%, and (iii) >=90% (Hint check out the substr() function in base R and the case_when() function from dplyr – see example below).
summary_percent_cocoa <- chocolate %>%
  select(country_of_bean_origin, cocoa_percent, rating) %>%
  filter(country_of_bean_origin %in% countries_with_enough_reviews) %>%
  mutate(numeric_cocoa_percent = as.numeric(sub("%", "", cocoa_percent))) %>%
  mutate(cocoa_percent_category = case_when(
    numeric_cocoa_percent < 60 ~ "<60%",
    60 <= numeric_cocoa_percent &  numeric_cocoa_percent < 70 ~ ">=60 to <70%",
    70 <= numeric_cocoa_percent & numeric_cocoa_percent < 90 ~ ">=70 to <90%",
    numeric_cocoa_percent >= 90 ~ ">=90%"
  )) %>%
  mutate(cocoa_percent_category = as.factor(cocoa_percent_category))
  
# third step: Using the new column described in #2, re-order the factor levels (if needed) to be starting with the smallest percentage group and increasing to the largest percentage group (Hint check out the fct_relevel() function from forcats).
summary_percent_cocoa %>%
  arrange(cocoa_percent_category)

# fourth step: For each country, make a set of four side-by-side boxplots plotting the groups on the x-axis and the ratings on the y-axis. These plots should be faceted by country.
ggplot(summary_percent_cocoa, aes(x = cocoa_percent_category , y = rating)) + 
  geom_line() + 
  #facet_grid(rows=vars(country_of_bean_origin),scales="free") +
  labs(x = "countries", y = "ratings", title = "Ratings by Chocolate Percentage Category in Each Country") +
  theme(
    plot.title = element_text(hjust = 0.5)  # Center the title
  )
```

## Part 2: Join two datasets together

```{r}
continent_data <- as_tibble(gapminder) %>%
  select(country, continent) %>%
  rename("country_of_bean_origin" = "country") %>%
  unique()
  
chocolate_w_continent <- chocolate %>%
  left_join(continent_data, by = "country_of_bean_origin")
```